Parenting Observational Study E2

### 2019-03-14 paper/cogsci2019/cogsci2019.Rmd
clean version

### Participants. 
`r nrow(d)` infants (F = 42, M = 18) aged 6-24 months (20 6-11.9 month-olds, 20 12-17.9 month-olds, and 20 18-24 month-olds) and their parents participated in a museum in northern California. 
We included infants who were exposed to English at least 50 percent of the time (n = 58) or who were exposed less but whose participating parent reported that they primarily speak English with their child at home (n = 2). 
Sixty-one% of participants (n = 37) had been exposed to two or more languages as indicated by their parent. 
Parents identified their children as White (n = 25), Asian (n = 11), African American/Black (n = 2), Biracial (n = 12), other (n = 5), or declined to state (n = 5). 
Fifteen parents reported their child was of Hispanic origin. 
Parents tended to be highly-educated, with reports of highest level of education ranging from completed high school (n = 5), some college (n = 6), four-year college (n = 14), some graduate school (n = 1), to completed graduate school (n = 28) or declined to state (n = 6).

@@

### Joint Attention Coding Procedure.
The video of each session was manually coded for episodes of joint attention using the Datavyu software [@datavyu]. 
The video taken at floor level was coded by default, but the other video was referred to if the participants were not visible or if there was technical difficulty with the first camera. 
Each session's video was coded for episodes of coordinated joint attention, episodes of passive joint attention, and parental bids for joint attention. 
Parental bids for joint attention were defined as any attempt to initiate joint attention (i.e labeling, pointing, or otherwise drawing attention to an object) that did not result in passive or coordinated joint attention. 
If more than 3 seconds elapsed between bids, they were coded as separate attempts. 
An episode of joint attention was considered passive if both participants visually focused on an object for a minimum of 3 seconds but the child did not acknowledge the parent. 
If either participant looked away from the object for less than 3 seconds and then returned to the same object it was considered part of the same period of joint attention. 
Episodes joint attention were considered coordinated if both participants visually focused on an object for a minimum of 3 seconds and at some point in the interaction the child indicated awareness of interaction with some overt behavior toward the parent. 
This could be looks to the parent's face, gestures, vocalizations, or turn-taking. 
If either participant looked away for less than 3 seconds and then returned to the same object, it was coded as part of the same period of joint attention. 
A second coder independently coded a third of the videos (i.e., 20 of the 60 videos, approximately equally distributed across ages) to establish reliability. 
The two coders had a reliability of ICC = 0.80 with 95% confident interval (CI) = [0.57,0.92] (p < 0.05) for number of parent bids for JA; ICC = 0.20 with 95% CI = [-0.26,0.58] for number of passive JA episodes; ICC = 0.66 with 95% CI = [0.32,0.85] (p < 0.05) for number of coordinated JA episodes; ICC = 0.24 with 95% CI = [-0.21,0.61] for total duration of passive JA episodes, and ICC = 0.62 with 95% CI = [0.27,0.83] (p < 0.05) for total duration of coordinated JA episodes.

@@

```{r, e1-lexdiv-regressions, echo=F, include=F}
# Predicting lexical diversity based on experimental condition, PAQ, demographics. 
modTTR <- lmer(TTR ~ condition + age + gender + parent_ed + (1|video), data=lmer_data)
summary(modTTR) #

modMTLD <- lmer(MTLD ~ condition + age + gender + parent_ed + (1|video), data=lmer_data)
summary(modMTLD)

# predict word tokens
tokens_mod <- lmer(tokens ~ condition*EL + condition*AA + condition*RR + age + gender + parent_ed + (1|video),  data=lmer_data)
summary(tokens_mod) # condition *

# predict word types
types_mod <- lmer(types ~ condition*EL + condition*AA + condition*RR  + age + gender + parent_ed + (1|video), data=lmer_data)
summary(types_mod)
```

@@

# Total number of bids
bids_mod <- lmer(bids_tot ~ Condition *  EL  + Condition * AA + Condition * RR + age + gender + parent_ed +
                           (1| Video), data = lmer_data)
summary(bids_mod)

# Episodes of coordinated joint attention.
cja_mod <- lmer(cja ~ Condition * EL + Condition * AA + Condition * RR  + age + gender + parent_ed +
                           (1| Video), data = lmer_data)
summary(cja_mod)

# Episodes of passive joint attention.
pja_mod <- lmer(pja ~ Condition * EL + Condition * AA + Condition * RR  + age + gender + parent_ed +
                           (1| Video), data = lmer_data)
summary(pja_mod)

# Total duration of passive joint attention.
pja_time_mod <- lmer(pja_length ~ Condition * EL + Condition * AA + Condition * RR  + age + gender + parent_ed +
                           (1| Video), data = lmer_data)
summary(pja_time_mod)

# Total duration of coordinated joint attention.
cja_time_mod <- lmer(cja_length ~ Condition * EL + Condition * AA + Condition * RR  + age + gender + parent_ed +
                           (1| Video), data = lmer_data)
summary(cja_time_mod)
```

@@

There is (was?) a main effect of condition on total bids for joint attention. Parents in the experimental condition (i.e., those who saw a video demonstrating an activity) made a greater number of bids for joint attention with their child.
There was no effect of condition on the number of episodes of either passive or coordinated joint attention, or the duration of these episodes. 
There is a marginal effect of gender on bids for joint attention, with parents of males producing more bids. There is a marginal interaction between RR scores and condition on passive joint attention, such that the experimental condition increased the number of episodes of PJA to a greater extent for people with high RR scores.
While the electronically-delivered parenting advice increased the number of bids for joint attention by parents, it did not significantly effect the number or duration of episodes of joint attention. One possibility is that child variables had a comparatively larger impact on the attainment of joint attention.

@@

## Lexical Diversity
```{r, e2-lexdiv-regressions, echo=F, include=F}
# Predicting lexical diversity based on experimental condition. 
modTTR <- lmer(TTR ~ age * condition + (1 | video), data=lmer_data) 
summary(modTTR)

modMTLD <- lmer(MTLD ~ age * condition + (1 | video), data=lmer_data)
summary(modMTLD)

# predict word tokens
tokens_mod <- lmer(tokens ~ age * condition + (1 | video), data=lmer_data)
summary(tokens_mod)

# predict word types
types_mod <- lmer(types ~ age * condition + (1 | video), data=lmer_data)
summary(types_mod)
```

We fit a mixed-effects linear regression predicting TTR and MTLD as a function of age (scaled and 0-centered) and condition with a random intercept per video using lme4 [@lme4].
There was significantly lower TTR in the Video condition (mean: 0.38) than in the Control condition (mean: 0.47, $\beta=-.09$, t(8) = 3.16, *p*<.05). 
There was no significant effect of age.
A similar mixed-effects linear regression instead predicting MTLD found no significant effects of age or condition.
Figure 1 shows the mean of each lexical diversity measure (TTR and MTLD) by condition.

@@

## Joint Attention
```{r, e2ja-regressions, echo=F, include=F}
x<- load(paste(e2path,"joint_attention/Exp2_joint_attention_data.RData",sep=''))
d$Condition <- factor(d$condition, levels = c("con","exp"), labels = c("Control","Video"))

lmer_data <- d %>%
  mutate(Condition = factor(Condition), 
         bids_tot = as.numeric(bids),
         Video = as.factor(video))

# Total number of bids
bids_mod <- lmer(bids_tot ~ age * condition + (1 | video), data = lmer_data)
summary(bids_mod)

# Episodes of coordinated joint attention.
cja_mod <- lmer(cja ~ age * condition + (1 | video), data = lmer_data)
summary(cja_mod)

# Episodes of passive joint attention.
pja_mod <- lmer(pja ~ age * condition + (1 | video), data = lmer_data)
summary(pja_mod)

# Total duration of passive joint attention.
pja_time_mod <- lmer(pja_length ~ age * condition + (1 | video), data = lmer_data)
summary(pja_time_mod)

# Total duration of coordinated joint attention.
cja_time_mod <- lmer(cja_length ~ age * condition + (1 | video), data = lmer_data)
summary(cja_time_mod)
```

>>>>

### Participants. 
`r nrow(d)` infants (F = 42, M = 18) aged 6-24 months (20 6-11.9 month-olds, 20 12-17.9 month-olds, and 20 18-24 month-olds) and their parents participated in a museum in northern California. 
We included infants who were exposed to English at least 50 percent of the time (n = 58) or who were exposed less but whose participating parent reported that they primarily speak English with their child at home (n = 2). 
61% of participants (n = 37) had been exposed to two or more languages as indicated by their parent. 
Parents identified their children as White (n = 25), Asian (n = 11), African American/Black (n = 2), Biracial (n = 12), other (n = 5), or declined to state (n = 5). 
Fifteen parents reported their child was of Hispanic origin. 
Parents tended to be highly-educated, with reports of highest level of education ranging from completed high school (n = 5), some college (n = 6), four-year college (n = 14), some graduate school (n = 1), to completed graduate school (n = 28) or declined to state (n = 6).

@@

### Joint Attention Coding Procedure.
The video of each session was manually coded for episodes of joint attention using the Datavyu software [@datavyu]. 
The video taken at floor level was coded by default, but the other video was referred to if the participants were not visible or if there was technical difficulty with the first camera. 
Each session's video was coded for episodes of coordinated joint attention, episodes of passive joint attention, and parental bids for joint attention. 
Parental bids for joint attention were defined as any attempt to initiate joint attention (i.e labeling, pointing, or otherwise drawing attention to an object) that did not result in passive or coordinated joint attention. 
If more than 3 seconds elapsed between bids, they were coded as separate attempts. 
An episode of joint attention was considered passive if both participants visually focused on an object for a minimum of 3 seconds but the child did not acknowledge the parent. 
If either participant looked away from the object for less than 3 seconds and then returned to the same object it was considered part of the same period of joint attention. 
A joint attention episode was considered coordinated if both participants visually focused on an object for a minimum of 3 seconds and at some point in the interaction the child indicated awareness of interaction with some overt behavior toward the parent such as looking at their face, gesturing, vocalizing, or turn-taking.

A second coder independently coded a third of the videos (i.e., 20 of the 60 videos, approximately equally distributed across ages) to establish reliability. 
The two coders had a reliability of ICC = 0.80 with 95% confident interval (CI) = [0.57,0.92] (p < 0.05) for number of parent bids for JA; ICC = 0.20 with 95% CI = [-0.26,0.58] for number of passive JA episodes; ICC = 0.66 with 95% CI = [0.32,0.85] (p < 0.05) for number of coordinated JA episodes; ICC = 0.24 with 95% CI = [-0.21,0.61] for total duration of passive JA episodes, and ICC = 0.62 with 95% CI = [0.27,0.83] (p < 0.05) for total duration of coordinated JA episodes.

@@

```{r, e1-lexdiv-regressions, echo=F, include=F}
# Predicting lexical diversity based on experimental condition, PAQ, demographics. 
modTTR <- lmer(TTR ~ condition + age + gender + parent_ed + (1|video), data=lmer_data)
summary(modTTR) # condition t(39.9) = 4.25 p<.001

modMTLD <- lmer(MTLD ~ condition + age + gender + parent_ed + (1|video), data=lmer_data)
summary(modMTLD) # condition t(43) = 4.69 p<.001

# predict word tokens
tokens_mod <- lmer(tokens ~ condition*EL + condition*AA + condition*RR + age + gender + parent_ed + (1|video),  data=lmer_data)
summary(tokens_mod) # condition t(34.4) = 2.19 p<.05

# predict word types
types_mod <- lmer(types ~ condition*EL + condition*AA + condition*RR  + age + gender + parent_ed + (1|video), data=lmer_data)
summary(types_mod) # nothing
```

@@

# Total number of bids
bids_mod <- lmer(bids_tot ~ Condition *  EL  + Condition * AA + Condition * RR + age + gender + parent_ed +
                           (1| Video), data = lmer_data)
summary(bids_mod) # condition beta = 3.51, t(40.3) = 2.95 p<.01

# Episodes of coordinated joint attention.
cja_mod <- lmer(cja ~ Condition * EL + Condition * AA + Condition * RR  + age + gender + parent_ed +
                           (1| Video), data = lmer_data)
summary(cja_mod) # NA

# Episodes of passive joint attention.
pja_mod <- lmer(pja ~ Condition * EL + Condition * AA + Condition * RR  + age + gender + parent_ed +
                           (1| Video), data = lmer_data)
summary(pja_mod) # marginal RR, beta=-1.07 p=.07, condition*RR beta=1.83 p<.05

# Total duration of passive joint attention.
pja_time_mod <- lmer(pja_length ~ Condition * EL + Condition * AA + Condition * RR  + age + gender + parent_ed +
                           (1| Video), data = lmer_data)
summary(pja_time_mod) # NA

# Total duration of coordinated joint attention.
cja_time_mod <- lmer(cja_length ~ Condition * EL + Condition * AA + Condition * RR  + age + gender + parent_ed +
                           (1| Video), data = lmer_data)
summary(cja_time_mod) # marginal parent education p=.09
```

@@

There was a main effect of condition on total bids for joint attention. Parents in the experimental condition (i.e., those who saw a video demonstrating an activity) made a greater number of bids for joint attention with their child.
There was no effect of condition on the number of episodes of either passive or coordinated joint attention, or the duration of these episodes. 
There is a marginal effect of gender on bids for joint attention, with parents of males producing more bids. There is a marginal interaction between RR scores and condition on passive joint attention, such that the experimental condition increased the number of episodes of PJA to a greater extent for people with high RR scores.
While the electronically-delivered parenting advice increased the number of bids for joint attention by parents, it did not significantly effect the number or duration of episodes of joint attention. One possibility is that child variables had a comparatively larger impact on the attainment of joint attention.

@@

## Lexical Diversity
```{r, e2-lexdiv-regressions, echo=F, include=F}
# Predicting lexical diversity based on experimental condition. 
modTTR <- lmer(TTR ~ age * condition + (1 | video), data=lmer_data) 
summary(modTTR) # condition beta=-.09  t(8.7) = 3.06 p=.01

modMTLD <- lmer(MTLD ~ age * condition + (1 | video), data=lmer_data)
summary(modMTLD) # NA

# predict word tokens
tokens_mod <- lmer(tokens ~ age * condition + (1 | video), data=lmer_data)
summary(tokens_mod) # NA

# predict word types
types_mod <- lmer(types ~ age * condition + (1 | video), data=lmer_data)
summary(types_mod) # NA
```

We fit a mixed-effects linear regression predicting TTR and MTLD as a function of age (scaled and 0-centered) and condition with a random intercept per video using lme4 [@lme4].
There was significantly lower TTR in the Video condition (mean: 0.38) than in the Control condition (mean: 0.47, $\beta=-.09$, t(8.7) = 3.06, *p*=.01).
There was no significant effect of age.
A similar mixed-effects linear regression instead predicting MTLD found no significant effects of age or condition.
Figure 1 shows the mean of each lexical diversity measure (TTR and MTLD) by condition.

@@

## Joint Attention
```{r, e2ja-regressions, echo=F, include=F}
load(paste(e2path,"joint_attention/Exp2_joint_attention_data.RData",sep=''))
d$Condition <- factor(d$condition, levels = c("con","exp"), labels = c("Control","Video"))
e2ja = d

t.test(subset(e2ja, Condition=="Control")$bids, subset(e2ja, Condition=="Video")$bids)

# Total number of bids
bids_mod <- lmer(bids ~ age * Condition + (1 | video), data = e2ja)
summary(bids_mod) # singular fit, covergence code: 0 - no effects

# Episodes of coordinated joint attention.
cja_mod <- lmer(cja ~ age * condition + (1 | video), data = e2ja)
summary(cja_mod) # marginal condition: t(79) = 1.88 beta=-5.85 p=.06, age*condition beta=4.92 t(79) = 2.35 p=.02
# also singular fit, covergence code: 0

# Episodes of passive joint attention.
pja_mod <- lmer(pja ~ age * condition + (1 | video), data = e2ja)
summary(pja_mod) # age beta=-2.32 t(26.6) = 2.23 p<.05

# Total duration of passive joint attention.
pja_time_mod <- lmer(pja_length ~ age * condition + (1 | video), data = e2ja)
summary(pja_time_mod) # age beta=-15512 t(79) = 2.40 p<.05

# Total duration of coordinated joint attention.
cja_time_mod <- lmer(cja_length ~ age * condition + (1 | video), data = lmer_data)
summary(cja_time_mod) # marginal age: beta=31757 t(79) = 1.81 p=.07
```

### 2019-03-14 paper/cogsci2019/cogsci2019.tex
clean version

60 infants (F = 42, M = 18) aged 6-24 months (20 6-11.9 month-olds, 20
12-17.9 month-olds, and 20 18-24 month-olds) and their parents
participated in a museum in northern California. We included infants who
were exposed to English at least 50 percent of the time (n = 58) or who
were exposed less but whose participating parent reported that they
primarily speak English with their child at home (n = 2). Sixty-one\% of
participants (n = 37) had been exposed to two or more languages as
indicated by their parent. Parents identified their children as White (n
= 25), Asian (n = 11), African American/Black (n = 2), Biracial (n =
12), other (n = 5), or declined to state (n = 5). Fifteen parents
reported their child was of Hispanic origin. Parents tended to be
highly-educated, with reports of highest level of education ranging from
completed high school (n = 5), some college (n = 6), four-year college
(n = 14), some graduate school (n = 1), to completed graduate school (n
= 28) or declined to state (n = 6).

@@

84 infants (F = 37, M = 47) aged 12-24 months (42 12-17.9 month-olds, 42
18-24 month-olds) and their parents participated in the same museum as
Experiment 1. We included infants who were exposed to English at least
75 percent of the time or who were exposed less but whose participating
parent reported that they primarily speak English with their child at
home. Forty nine\% of participants (n = 41) had been exposed to two or
more languages as indicated by their parent. Parents identified their
children as White (n = 39), Asian (n = 20), African American/Black (n =
1), Biracial (n = 9), other (n = 7), or declined to state (n = 8).
Sixteen parents reported their child was of Hispanic origin. Parents
tended to be highly- educated, with reports of highest level of
education ranging from completed high school (n = 0), some college (n =
5), four-year college (n = 28), some graduate school (n = 2), to
completed graduate school (n = 35) or declined to state (n = 14).

@@

We fit a mixed-effects linear regression predicting TTR and MTLD as a
function of age (scaled and 0-centered) and condition with a random
intercept per video using lme4 (Bates et al., 2015). There was
significantly lower TTR in the Video condition (mean: 0.38) than in the
Control condition (mean: 0.47, \(\beta=-.09\), t(8) = 3.16,
\emph{p}\textless{}.05). There was no significant effect of age. A
similar mixed-effects linear regression instead predicting MTLD found no
significant effects of age or condition. Figure 1 shows the mean of each
lexical diversity measure (TTR and MTLD) by condition.

>>>>

60 infants (F = 42, M = 18) aged 6-24 months (20 6-11.9 month-olds, 20
12-17.9 month-olds, and 20 18-24 month-olds) and their parents
participated in a museum in northern California. We included infants who
were exposed to English at least 50 percent of the time (n = 58) or who
were exposed less but whose participating parent reported that they
primarily speak English with their child at home (n = 2). 61\% of
participants (n = 37) had been exposed to two or more languages as
indicated by their parent. Parents identified their children as White (n
= 25), Asian (n = 11), African American/Black (n = 2), Biracial (n =
12), other (n = 5), or declined to state (n = 5). Fifteen parents
reported their child was of Hispanic origin. Parents tended to be
highly-educated, with reports of highest level of education ranging from
completed high school (n = 5), some college (n = 6), four-year college
(n = 14), some graduate school (n = 1), to completed graduate school (n
= 28) or declined to state (n = 6).

@@

83 infants (F = 37, M = 47) aged 12-24 months (42 12-17.9 month-olds, 42
18-24 month-olds) and their parents participated in the same museum as
Experiment 1. We included infants who were exposed to English at least
75 percent of the time or who were exposed less but whose participating
parent reported that they primarily speak English with their child at
home. Forty nine\% of participants (n = 41) had been exposed to two or
more languages as indicated by their parent. Parents identified their
children as White (n = 39), Asian (n = 20), African American/Black (n =
1), Biracial (n = 9), other (n = 7), or declined to state (n = 8).
Sixteen parents reported their child was of Hispanic origin. Parents
tended to be highly- educated, with reports of highest level of
education ranging from completed high school (n = 0), some college (n =
5), four-year college (n = 28), some graduate school (n = 2), to
completed graduate school (n = 35) or declined to state (n = 14).

@@

We fit a mixed-effects linear regression predicting TTR and MTLD as a
function of age (scaled and 0-centered) and condition with a random
intercept per video using lme4 (Bates et al., 2015). There was
significantly lower TTR in the Video condition (mean: 0.38) than in the
Control condition (mean: 0.47, \(\beta=-.09\), t(8.7) = 3.06,
\emph{p}=.01). There was no significant effect of age. A similar
mixed-effects linear regression instead predicting MTLD found no
significant effects of age or condition. Figure 1 shows the mean of each
lexical diversity measure (TTR and MTLD) by condition.

### 2019-03-14 parenting_obs_e2/Lexical_Diversity_Analysis.Rmd
clean version

# age is "1" or "2", but should be continuous value (in mos)
d$age = as.numeric(d$age)
d$Condition <- factor(d$condition, levels=c("con","exp"), labels=c("Control","Video"))
save(d, file="Exp2_cleaned_data.Rdata")
```

>>>>

dem <- read.csv("anonymized_demographics_parenting_obs_e2.csv", header=T)
dem$age = as.numeric(as.character(dem$age))
d$age = NULL
dem$X = NULL
d <- d %>% left_join(dem, by=c("sid" = "SID")) %>%
  filter(age <= 2.0)

d$Condition <- factor(d$condition, levels=c("con","exp"), labels=c("Control","Video"))
save(d, file="Exp2_cleaned_data.Rdata")
```

### 2019-03-14 parenting_obs_e2/joint_attention/parenting_obs_JA.Rmd
clean version

Total number of bids (successful and unsuccessful)
```{r, width=4, height=4}
save(d, file="Exp2_joint_attention_data.RData")
ms_bids <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "bids")

>>>>

Total number of bids (successful and unsuccessful)
```{r, width=4, height=4}
dem <- read.csv("../anonymized_demographics_parenting_obs_e2.csv", header=T)
dem$age = as.numeric(as.character(dem$age))
d$age = NULL
d$X = NULL
d$X.1 = NULL
d$X.2 = NULL
dem$X = NULL
d <- d %>% left_join(dem, by=c("sid" = "SID")) %>%
  filter(age <= 2.0)

save(d, file="Exp2_joint_attention_data.RData")
ms_bids <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "bids")
